---
title: "FARS functions README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# farsfunctions

<!-- badges: start -->
[![R-CMD-check](https://github.com/hriisalu/farsfunctions/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hriisalu/farsfunctions/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Overview

The `"farsfunctions"` package provides functions to work with the [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars) data. FARS is a nationwide census providing annually updated data regarding fatal injuries suffered in motor vehicle traffic crashes in the United States. 

## Installation

You can install the development version of `"farsfunctions"` package from GitHub.

```{r, message=FALSE, warning=FALSE}
library(devtools)
devtools::install_github("hriisalu/farsfunctions", build_vignettes = TRUE, force = TRUE)
```

## Using

### Attaching the package

```{r}
library(farsfunctions)
```

### Reading FARS data

The `"fars_read"` function allows users to read FARS data from a csv file. It takes the path to the csv file as input and returns a tibble containing the data. If the file does not exist, an error message is displayed.

```{r}
# Example
fars_read("accident_2013.csv.bz2") %>% 
  head(5)
```

### Creating file name

The `"make_filename"` function generates the filename for a FARS data file based on the input year.

```{r}
# Example
make_filename(2013)
```

### Reading FARS data for multiple years

The `"fars_read_years"` function reads FARS data for multiple years and combines them into a tibble. It takes a vector of years as input and returns a list of tibbles, each containing data for a specific year.

```{r}
# Example
fars_read_years(c(2013, 2014)) %>% 
  lapply(head, n = 5)
```

### Summarizing FARS data for multiple years

The `"fars_summarize_years"` function summarizes FARS data for multiple years, counting accidents per month. It takes a vector of years as input and returns a tibble summarizing the number of accidents per month for each year.

```{r}
# Example
fars_summarize_years(c(2013, 2014)) %>% 
  head(5)
```


### Mapping FARS data for a specific state and year

The `"fars_summarize_years"` function shows FARS data on a map for a specific state and year. It takes the state number and year as input and returns a plot of FARS data on a map.

```{r}
# Example
fars_map_state(50, 2013)
```

For more detailed information on each function, refer to the function documentation.


## License

GPL

