---
title: "FARS functions README"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# farsfunctions

<!-- badges: start -->
[![R-CMD-check](https://github.com/hriisalu/farsfunctions/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hriisalu/farsfunctions/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## Note about the badge

The package uses the R-CMD-check badge from GitHub instead of Travis because Travis does not offer a free version.

## Overview

The `"farsfunctions"` package provides functions to work with the [Fatality Analysis Reporting System (FARS)](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars) data. FARS is a nationwide census providing annually updated data regarding fatal injuries suffered in motor vehicle traffic crashes in the United States. 

## Installation

You can install the development version of `"farsfunctions"` package from GitHub.

```{r, eval=FALSE}
library(devtools)
devtools::install_github("hriisalu/farsfunctions", build_vignettes = TRUE, force = TRUE)
```

## Usage

```{r}
# Attaching the package
library(farsfunctions)
```

```{r}
# Reading FARS data
fars_read("accident_2013.csv.bz2") %>% 
  head(5)

# Creating file name
make_filename(2013)

# Reading FARS data for multiple years
fars_read_years(c(2013, 2014)) %>% 
  lapply(head, n = 5)

# Summarizing FARS data for multiple years
fars_summarize_years(c(2013, 2014)) %>% 
  head(5)

# Mapping FARS data for a specific state and year
fars_map_state(50, 2013)

```

For more detailed information on each function, refer to the function documentation.

## License

GPL

