---
title: "fars_functions_vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fars_functions_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(farsfunctions)
```

## Introduction

The `"farsfunctions"` package provides functions to work with the Fatality Analysis Reporting System (FARS) data. FARS is a nationwide census providing annually updated data regarding fatal injuries suffered in motor vehicle traffic crashes in the United States. This vignette explains the functionalities provided by this package. 

## Reading FARS data

The `"fars_read"` function allows users to read FARS data from a csv file. It takes the path to the csv file as input and returns a tibble containing the data. If the file does not exist, an error message is displayed.

```{r echo = FALSE}
# Example
data_2013 <- fars_read("data/accident_2013.csv.bz2")
```

## Creating file name

The `"make_filename"` function generates the filename for a FARS data file based on the input year.

```{r echo = FALSE}
# Example
filename_2013 <- make_filename(2013)
```

## Reading FARS data for multiple years

The `"fars_read_years"` function reads FARS data for multiple years and combines them into a tibble. It takes a vector of years as input and returns a list of tibbles, each containing data for a specific year.

```{r echo = FALSE}
# Example
data_2013_2014 <- fars_read_years(c(2013, 2014))
```

## Summarizing FARS data for multiple years

The `"fars_summarize_years"` function summarizes FARS data for multiple years, counting accidents per month. It takes a vector of years as input and returns a tibble summarizing the number of accidents per month for each year.

```{r echo = FALSE}
# Example
summary_2013_2014 <- fars_summarize_years(c(2013, 2014))
```

## Mapping FARS data for a specific state and year

The `"fars_summarize_years"` function shows FARS data on a map for a specific state and year. It takes the state number and year as input and returns a plot of FARS data on a map.

```{r echo = FALSE}
# Example
map_2013 <- fars_map_state(50, 2013)
```
This vignette provides an overview of the functionalities offered by the `"farsfunctions"` package. For more detailed information on each function, refer to the function documentation.

